My title

### 2022-02-28 outputs/paper/paper.Rmd
fixed conflict, add abstract

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Lantao Cui
  - Huakun Shen
  - Sijian Wu
thanks: "Replication avaliable at https://doi.org/10.48152/ssrp-mbxx-jm47. Code and data avaliable at https://github.com/cuilantao/Investigation-of-falling-US-Birth-Rates-since-the-great-recession"         
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This paper is based on Kearney's paper The Puzzle of Falling US Birth Rates since the Great Recession Between 1980. ^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44. However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years. In this paper, we will expand our research on multiple countries and find out whether the level of impact from recession would be correlated with the changes in the birth rate for different countries. We will also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examine its effect"
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
header-includes:
  - \usepackage{subfig}
---

@@

The Great Recession has been a world-wide crisis that reconstructed the economy, and was deep, synchronous and global. Canadian GDP dropped by 3.3 percent, US GDP by close to 4 percent, that of the Euro area by 5 percent and that of Japan by more than 8 percent. The impact on international trade has been severe that all G20 countries experienced serious reductions in trade[@recession]. Apart from economic impact, the great recession also had a negative impact on country's birth rate, Figure 1 shows the US birth rate from 1980 - 2020, as we can see, the birth rate declined significantly after 2007, which is the time of the Great Recession.

```{r, fig.cap="US Birth Rate in the past 40 years", echo=FALSE, warning=FALSE, message = FALSE, fig.width=10, fig.height=4}
US_birth_date_from_1985_to_2020 <- read_csv(here::here("inputs/data/Birthrate.csv"))

In the United States, unemployment rates, mortgage foreclosures, and poverty rates rose while housing values fell, but the extent of these changes varied widely across local areas. Another impact from depressed economic condition is delayed childbearing, and others to hasten it. The fertility rate declined at the national level, dropping from a recent high in 2007 of 69.5 births per 1,000 women aged 15 to 44, to 63.2 for 2012.1 There was, however, great variation by state, age, and ethnicity, with younger and Hispanic women showing disproportionate decreases[@fertility]. The significant consequence on fertility rate would make us wonder how severe is America being affected by the great recession. However,  by comparing currency devaluations, equity market declines, and rising sovereign bond spreads, the U.S. is the third leas affected country[@rank].

@@

As mentioned by @cite_original, the basic approach to many economic problem is maximizing utility. This approach recognized that children bring people utility but also with associated costs including time and money. This approach to modeling the decision to have a child leads to standard predictions of price and income effects. Direct consideration of costs could be like housing and medical care. Other times the price effect is indirect, like the opportunity cost of a woman’s time. Holding prices and quality constant, an increase in income will lead people to choose to have more children. This positive relationship between income and births may come as a surprise to some readers, given the negative relationship between birth rates and income or per capita GDP observed over time and place[@cite_original]. However, this is enough for us to consider whether the birth rates are world-wide affected in the same negative direction after the Great Recession.

@@

## Data Source
We generated data from U.S. Birth Rate 1950-2022[@birthrate]. We selected top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank].

@@

```{r,echo=FALSE}
#Generating time series graphs of births per 1000 people for each least affected country

@@

```{r, echo=FALSE}
#Generating time series graphs of births per 1000 people for each most affected country

@@

# Model

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## LGBT Section

>>>>

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Lantao Cui
  - Huakun Shen
  - Sijian Wu
thanks: "Replication avaliable at https://doi.org/10.48152/ssrp-mbxx-jm47. Code and data avaliable at https://github.com/cuilantao/Investigation-of-falling-US-Birth-Rates-since-the-great-recession"         
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This paper is based on Kearney's paper \"The Puzzle of Falling US Birth Rates since the Great Recession\". ^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, US birth rate varied between a tight range of roughly 65 to 70 birth per 1000 women between the age of 15 to 44. However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years. In this paper, we will expand our research on multiple countries and find out whether the level of impact from recession would be correlated with the changes in the birth rate for different countries. We will also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examine its effect"
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
header-includes:
  - \usepackage{subfig}
---

@@

The Great Recession has been a world-wide crisis that reconstructed the economy, and was deep, synchronous and global. Canadian GDP dropped by 3.3 percent, US GDP by close to 4 percent, that of the Euro area by 5 percent and that of Japan by more than 8 percent. The impact on international trade has been severe that all G20 countries experienced serious reductions in trade[@recession]. Apart from economic impact, the great recession also had a negative impact on country's birth rate.

In the United States, unemployment rates, mortgage foreclosures, and poverty rates rose while housing values fell, but the extent of these changes varied widely across local areas. Another impact from depressed economic condition is delayed childbearing, and others to hasten it. The fertility rate declined at the national level, dropping from a recent high in 2007 of 69.5 births per 1,000 women aged 15 to 44, to 63.2 for 2012.1 There was, however, great variation by state, age, and ethnicity, with younger and Hispanic women showing disproportionate decreases[@fertility]. The significant consequence on fertility rate would make us wonder how severe is America being affected by the great recession. However,  by comparing currency devaluations, equity market declines, and rising sovereign bond spreads, the U.S. is the third less affected country[@rank].

@@

As mentioned by @cite_original, the basic approach to many economic problem is maximizing utility. This approach recognized that children bring people utility but also with associated costs including time and money. This approach to modeling the decision to have a child leads to standard predictions of price and income effects. Direct consideration of costs could be like housing and medical care. Other times the price effect is indirect, like the opportunity cost of a woman’s time. Holding prices and quality constant, an increase in income will lead people to choose to have more children. This positive relationship between income and births may come as a surprise to some readers, given the negative relationship between birth rates and income or per-capita GDP observed over time and place[@cite_original]. However, this is enough for us to consider whether the birth rates are world-wide affected in the same negative direction after the Great Recession.

@@

## Data Source
Part of our data are from the NCHS(National Center for Health Statistics)[@nvss], in particular, we acquired US birth rate data for the past 40 years, Figure \@ref(fig:fig-past-40-years) shows the US birth rate from 1980 - 2020, as we can see, the birth rate declined significantly after 2007, which is the time of the Great Recession. We also generated data from U.S. Birth Rate 1950-2022[@birthrate]. We selected top 3 countries that are affected most and least by the Great Recession according to rankings determined by comparing currency devaluations, equity market declines, and rising sovereign bond spreads These measures tend to track developments in the real economy during times of economic crisis when financial strain handicaps consumption, investment, and in many cases government spending, which limits GDP and employment growth[@rank]. TODO: ADD DATA SOURCE AND BREIFLY TALK ABOUT IT

## Data Characteristics

```{r, fig-past-40-years, fig.cap="US Birth Rate in the past 40 years", echo=FALSE, warning=FALSE, message = FALSE, fig.width=10, fig.height=4}
US_birth_date_from_1985_to_2020 <- read_csv(here::here("inputs/data/Birthrate.csv"))

@@

## Global Comparison of Fertility Rate Under The Great Recession
It shows that although the United States is at the epicenter of the global economic crisis, it is one of the countries least affected by the financial fallout. Large industrialized nations like the United States, Japan, and Germany have benefited from increasing global demand for relatively stable economies in which to invest. Instead, it is several developing countries, notably those with vulnerable capital accounts and weak macroeconomic fundamentals, that are experiencing severe economic downturns disproportionate to their roles in the crisis[@rank].

```{r, least-affected, echo=FALSE, fig.cap="Births per 1000 People of Top Least Affected Countries"}
#Generating time series graphs of births per 1000 people for each least affected country

@@

Figure \@ref(fig:least-affected) shows the changes of births per 1000 people of China, Japan, the U.S, and South Africa since 1950 with prediction for the future. The births rates of China and South Africa were obviously higher than the U.S. and Japan. However, around the year of recession, the birth rate of China and Japan dropped even harder than the U.S. on the absolute level that are lower than 10 births per 1000 people.

```{r, most-affected, echo=FALSE, fig.cap="Births per 1000 People of Top Most Affected Countries"}
#Generating time series graphs of births per 1000 people for each most affected country

@@

Figure \@ref(fig:most-affected) shows the changes of births per 1000 people of Ukraine, Argentina, Hungary, Poland since 1950 with prediction for the future. All these most affected countries have lower initial births rate than those least affected countries. However, the levels of births per 1000 people were similar to those least affected countries around recession except Argentina. Ukraine, Hungary, and Poland also showed a potential increase after recession but continued to decline afterwards.

We do observe declinations of world-wide birth rate despite of the impact level from recession. From the economic perspective, the countries who are least affected tend to have a stronger resistance to recession but not necessarily developed countries. While the most affected countries are mostly developing countries, they experienced a more significant plummet as percentage. It is also intriguing that the long-term births per 1000 people tend to converge to 10 for all countries mentioned above. However, the countries have higher resistance tend to delay the convergence which may derived from stronger economic status. Overall, the Great Recession was a turning point not only for the U.S. but also the whole world in the context of birth rate. The Great Recession may not be the direct cause of the world-wide declination in birth rates but it indeed provided an opportunity and provoked the desire for people to reconsider personal priorities with changed perspective of the society.

## Effect of LGBT movement
